m0.0.1: Microxanox - ...

### 2022-03-07 inst/manuscript/manuscript_microxanox.Rmd
Next Steps Manuscript

The aims of the `microxanox` package are twofold. Firstly, the package aims at reproducing the results shown by [@Bush2017]. To take one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED].

@@

We extended the model and added functionality for:

We will go through these additions in detail below.

@@

The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the odeSolve package [@REF_NEEDED]

@@

@@

The package will be updated based on further research activities. It will be tested continuously for the last revent, current and devel version of R.
Issues, questions and suggestions should be communicated through the issue tracker at [https://github.com/UZH-PEG/microxanox/issues](https://github.com/UZH-PEG/microxanox/issues)

# Usage of the Package
The package includes a full documentetion, which includes

1. Standard R help for each function (`?COMMAND` in R)
2. Two vignettes acompanying the paclage, also available under RUniverse [Vignette 1](LINK NEEDED)
3. This paper as a third vignette, also available under RUniverse [Vignette 1](LINK NEEDED)

## Simulation Preparation
1. Define a parameter set of class `runsim_parameter` (i.e. parameters and initial conditions) by using the functions `new_runsim_parameter()`, `new_initial_state()` and `new_strain_parameter()`.

- Set parameter
	- `new_xxx_parameter()`

## Simulation steps
2. The function call `run_simulation(parameter)` which will run a simulation using the parameter as defined in the object `parameter`.

- single simulation `run_simulation()`
- Stable State Finding 
	- Replication Method `run_replication_ssfind()`
	- Temporal Method `run_replication_ssfind()`
- {>>I would leave the `run_temporal_experiment()` out as it has no counterpart for the replication method <<}

>>>>

The microxanox package is a package for simulating a three functional group system (CB: cyanobacteria, PB: phototrophic sulfur bacteria, and SB: sulfate-reducing bacteria) with four chemical substrates (P: phosphorus, O: oxygen, SR: reduced sulfur, SO: oxidized sulfur). It includes feddback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).

The aims of the `microxanox` package are twofold. Firstly, the package aims at reproducing the results shown by [@Bush2017]. Secondly, to take these results one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED]. We extended the model and added functionality for:

@@

@@

The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the `deSolve` package [@Soetaert2010]

@@

For an example how these can be used see [@REF_NEEDED].

@@

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. {>>If we wnat that, we havt o include some, even rudimentary, GITHUB tests.<<}

Issues, questions and suggestions should be communicated through the issue tracker at [https://github.com/UZH-PEG/microxanox/issues](https://github.com/UZH-PEG/microxanox/issues)

# The Package
The package includes a full documentetion, which includes

1. Standard R help for each function (`?COMMAND` in R)
2. Two vignettes acompanying the package, also available under RUniverse [Vignette 1](LINK NEEDED) and [Vignette 2](LINK NEEDED)
3. This paper as a third vignette, also available under RUniverse [Vignette 3](LINK NEEDED)

The simulations can be split into different levels:

## The individuel simulation
The individual simulation (`run_simulation()` function) is the working horse in this package. In this function, the ODEs are solved. The function needs only one parameter - an object as created by the function `new_runsim_parameter()`. One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at [User Guide](Link Needed)

After the have been define, they can be used in the `run_simulation()` function. The function returns an object which is identical to the parameter, except of an additional slot, the results. This makes it possible to run the simulation again from the results from a previous simulation.

## Find a Steady State of a model
There are two methods for finding steady states. The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*). The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*).

### Replication Method
The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.

### Temporal Method
The temporal method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is *increased* in a step-wise fashion. In the other it is *decreased* in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step. We hereafter term this "the temporal method".
This is implemented in the function `run_temporal_ssfind()` while the parameter object is created by the function `new_temporal_ssfind_parameter()()`.'

For a more detailed walk-through of these two approaches and explanation please see the [User Guide](LINK NEEDED).

## Extract Stability Measures
From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`. These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change.
